---
title: "Secondary analysis - comparing the two approaches (inter-treatment vs. treatment vs. controls)"
author: "Francisca Rodriguez"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Data input**
```{r}
#load libraries 
library(phyloseq)
library(tibble)
library(dplyr)

#Load seqtable.nochim and taxa from dada2 files from Kacie
seqtable.nochim <- readRDS('./raw_data/dada2_seqtable.RDS')
taxa <- readRDS('./raw_data/RDP_Taxonomy_from_dada2.RDS')
strawb_miseq_map <- read.csv("./raw_data/Strawberry_Miseq_Map_forQIIME.csv", fileEncoding="UTF-8-BOM")

# add empty column
strawb_miseq_map <- add_column(strawb_miseq_map, SampleType = NA)

# get rid of unneeded cols
strawb_miseq_map <- select(strawb_miseq_map, 
                           SampleID, 
                           TimePoint, 
                           Type, 
                           PlantID, 
                           WeightGrams, 
                           SampleNotes, 
                           Water, 
                           Inoculation, 
                           Replicate, 
                           PlantNotes, 
                           SampleType)

# label sample types *INOCULATION 6 INCLUDED IN POST_INOC_POST_STRESS*
strawb_miseq_map$SampleType[strawb_miseq_map$TimePoint == "T0"] <- "post_inoc_pre_stress"
strawb_miseq_map$SampleType[strawb_miseq_map$TimePoint == "T1"] <- "post_inoc_post_stress"
strawb_miseq_map$SampleType[strawb_miseq_map$SampleNotes == "Empty Extraction Tube Control"] <- "DNA_neg"
strawb_miseq_map$SampleType[grep("Inoc", strawb_miseq_map$SampleNotes)] <- "slurry"
strawb_miseq_map$SampleType[grep("Inocolum 6", strawb_miseq_map$SampleNotes)] <- "di_water"
strawb_miseq_map$SampleType[grep("pre-inoc", strawb_miseq_map$SampleNotes)] <- "pre_inoc"

# add empty column for control
strawb_miseq_map <- add_column(strawb_miseq_map, Sample_or_Control = "True Sample")

#label controls (i6, DI water, DNA/PCR negative, pre-inoculated plants)
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$SampleType == "di_water"] <- "Control"
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$SampleType == "DNA_neg"] <- "Control"
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$SampleType == "pre_inoc"] <- "Control"
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$Inoculation == "6" & strawb_miseq_map$TimePoint == "T1"] <- "Control"

# remove unwanted samples
strawb_miseq_map <- strawb_miseq_map[grep("CSIA", strawb_miseq_map$SampleNotes, invert = TRUE),]
strawb_miseq_map <- strawb_miseq_map[strawb_miseq_map$Type != "Palmyra",]

#save new map as a .csv file
write.csv(strawb_miseq_map, "./combined_treatment_data/strawb_miseq_map2.csv")
```

**Phyloseq**
```{r}
# Hand off to Phyloseq ####
otu <- otu_table(seqtable.nochim,taxa_are_rows = FALSE)
tax <- tax_table(taxa)
met <- sample_data(strawb_miseq_map)
row.names(met) <- strawb_miseq_map$SampleID

ps2 <- phyloseq(otu,met,tax)

# generate the ASV0001, ASV0002 names
pretty_names <- paste("ASV", sprintf('%0.4d', 1:length(taxa_names(ps2))), sep = "")

# reassign the taxa_names from the sequences to the generated ASV names
taxa_names(ps2) <- pretty_names

#save phyloseq object
saveRDS(ps2, "./combined_treatment_data/ps2.rds")

```

**Decontaminate controls**
```{r}
library(decontam)

#remove taxa with only one read 
ps2 <- prune_taxa(taxa_sums(ps2) > 1, ps2)

#label controls (extraction negatives) and clean
sample_data(ps2)$is.neg <- sample_data(ps2)$Sample_or_Control == "Control"
contams <- isContaminant(ps2, method="prevalence", neg="is.neg")
table(contams$contaminant)

#Remove contaminants
ps.noncontam2 <- prune_taxa(!contams$contaminant, ps2)

#save phyloseq object
saveRDS(ps.noncontam2, "./combined_treatment_data/ps.noncontam2.rds")
```

**Subset by post-inoc post-stress**
```{r}
#post_inoc_post_stress table 
post_inoc_post_stress_ps <- subset_samples(ps.noncontam2, sample_data(ps.noncontam2)$SampleType=="post_inoc_post_stress")
post_inoc_post_stress_ps <- prune_taxa(taxa_sums(post_inoc_post_stress_ps) > 0, post_inoc_post_stress_ps)
post_inoc_post_stress_asv <- as.data.frame(otu_table(post_inoc_post_stress_ps))

#Hellinger transformation for all replicates in post-inoculation/post-stress ASV contingency table
transformed_pips_ASV <- hellinger(post_inoc_post_stress_asv)

#On this transformed data, calculate a “relative abundance” by summing the new totals for all ASVs (the transformed number of total reads) within a replicate and then dividing the transformed number of reads for each ASV by this total
rel_abund <- transformed_pips_ASV / rowSums(transformed_pips_ASV) 

#pull out tax table and metadata from post_inoc_post_stress_ps to create new phyloseq object
pips_metadata <- as(sample_data(post_inoc_post_stress_ps), 'matrix')
pips_metadata <- as.data.frame(pips_metadata)
pips_taxtable <- as(tax_table(post_inoc_post_stress_ps), 'matrix')

#Create new phyloseq object with transformed table
otu <- otu_table(rel_abund,taxa_are_rows = FALSE)
tax <- tax_table(pips_taxtable)
met <- sample_data(pips_metadata)
row.names(met) <- pips_metadata$SampleID

ps.pips2 <- phyloseq(otu,met,tax)

#Using this new table, calculate Bray-Cutis community dissimilarities for all replicates and build an NMDS visualization where you code inoculum type by color and water treatment by shape.
GP.ord <- ordinate(ps.pips2, "NMDS", "bray")
p1 = plot_ordination(ps.pips2, GP.ord, type="samples", color="Inoculation", shape = "Water", title="Bray-Cutis community dissimilarities for all replicates")
print(p1)


#save phyloseq object
saveRDS(ps.pips2, "./combined_treatment_data/ps.pips2.rds")
```

**Box and whisker plot of average richness**
```{r}
#Sum ASVs across samples to add to metadata
richness_sums <- rowSums(post_inoc_post_stress_asv)

#add column to new df with corresponding richness values
pips_complete2 <- dplyr::bind_cols(pips_metadata, richness_sums)

#rename column
names(pips_complete2)[14] <- 'SpeciesRichness'
str(pips_complete2)

#Grouped boxplot by inoculation and water treatment
ggplot(pips_complete2, aes(x=Inoculation, y=SpeciesRichness, fill=Water)) + 
    geom_boxplot() + theme_light()

#Grouped boxplot by inoculation and plant part
ggplot(pips_complete2, aes(x=Inoculation, y=SpeciesRichness, fill=Water)) + 
    geom_boxplot() + theme_light() + facet_wrap(~ Type)

#write this to .csv
write.csv(pips_complete2, "./combined_treatment_data/pips_complete2.csv")

```

**Outlier analysis: **
```{r}
#load libraries
library(tidyverse)
library(car)
library(ggpubr)
library(effects)
library(emmeans)

dat = read.csv("./combined_treatment_data/pips_complete2_edited.csv") |>
  mutate(
    Inoculation = as.character(Inoculation),
    Water = as.character(Water)
  )

# interaction model
fit1 = aov(log(SpeciesRichness) ~ Type + Inoculation * Water, data = dat)
Anova(fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate

# additive model
fit2 = aov(log(SpeciesRichness) ~ Type + Inoculation + Water, data = dat)
Anova(fit2, type = "2") # if you don't have interactions, type-2 sums-of-squares is appropriate

#plot to check residuals
plot(fit1)
plot(fit2)

# Studentized residuals
MASS::studres(fit1) |>
  abs() |>
  sort()

#remove the outliers, re-run the analysis
filtered_dat = dat[-c(88, 34, 91, 464, 133), ]

# interaction model
fit1 = aov(log(SpeciesRichness) ~ Type + Inoculation * Water, data = filtered_dat)
Anova(fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate

# additive model
fit2 = aov(log(SpeciesRichness) ~ Type + Inoculation + Water, data = filtered_dat)
Anova(fit2, type = "2") # if you don't have interactions, type-2 sums-of-squares is appropriate

#plot to check residuals
plot(fit1)
plot(fit2)

#Re-plot grouped boxplot by inoculation and plant part without the outliers
ggplot(filtered_dat, aes(x=Inoculation, y=log(SpeciesRichness), fill=Water)) + 
    geom_boxplot() + theme_light() + facet_wrap(~ Type)

```

**ANOVA: roots**
```{r}
root_dat = subset(filtered_dat, Type == "Root")

#interaction model
richness_root_fit1 = aov(log(SpeciesRichness) ~ InoculationLabeled * WaterLabeled, data = root_dat)
Anova(richness_root_fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(richness_root_fit1)
plot(allEffects(richness_root_fit1))

#additive model 
richness_root_fit2 = aov(log(SpeciesRichness) ~ InoculationLabeled + WaterLabeled, data = root_dat)
Anova(richness_root_fit2, type = "2")
plot(richness_root_fit2)
plot(allEffects(richness_root_fit2))
```

**Post-hoc test roots: planned comparisons**
```{r}
## Type III ANOVA ##

#apply the emmeans() command to the ANOVA object.
RichRootsW1<- emmeans(richness_root_fit1, ~ WaterLabeled)
RichRootsW1

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedW1 <- contrast(RichRootsW1, adjust = "none")
RootsPlannedW1

#apply the emmeans() command to the ANOVA object.
RichRootsI1<- emmeans(richness_root_fit1, ~ InoculationLabeled)
RichRootsI1

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedI1 <- contrast(RichRootsI1, adjust = "none")
RootsPlannedI1

## Type II ANOVA ##
#apply the emmeans() command to the ANOVA object.
RichRootsW2<- emmeans(richness_root_fit2, ~ WaterLabeled)
RichRootsW2

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedW2 <- contrast(RichRootsW2, adjust = "none")
RootsPlannedW2

#apply the emmeans() command to the ANOVA object.
RichRootsI2<- emmeans(richness_root_fit2, ~ InoculationLabeled)
RichRootsI2

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedI2 <- contrast(RichRootsI2, adjust = "none")
RootsPlannedI2
```

**ANOVA: shoots**
```{r}
shoot_dat = subset(filtered_dat, Type == "Shoot")

#interaction model
richness_shoot_fit1 = aov(log(SpeciesRichness) ~ InoculationLabeled * WaterLabeled, data = shoot_dat)
Anova(richness_shoot_fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(richness_shoot_fit1)
plot(allEffects(richness_shoot_fit1))

#additive model 
richness_shoot_fit2 = aov(log(SpeciesRichness) ~ InoculationLabeled + WaterLabeled, data = shoot_dat)
Anova(richness_shoot_fit2, type = "2")
plot(richness_shoot_fit2)
plot(allEffects(richness_shoot_fit2))
```

**Post-hoc test shoots: planned comparisons**
```{r}
## Type III ANOVA ##

#apply the emmeans() command to the ANOVA object.
RichShootsW1<- emmeans(richness_shoot_fit1, ~ WaterLabeled)
RichShootsW1

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedW1 <- contrast(RichShootsW1, adjust = "none")
ShootsPlannedW1

#apply the emmeans() command to the ANOVA object.
RichShootsI1<- emmeans(richness_shoot_fit1, ~ InoculationLabeled)
RichShootsI1

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedI1 <- contrast(RichShootsI1, adjust = "none")
ShootsPlannedI1

## Type II ANOVA ##
#apply the emmeans() command to the ANOVA object.
RichShootsW2<- emmeans(richness_shoot_fit2, ~ WaterLabeled)
RichShootsW2

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedW2 <- contrast(RichShootsW2, adjust = "none")
ShootsPlannedW2

#apply the emmeans() command to the ANOVA object.
RichShootsI2 <- emmeans(richness_shoot_fit2, ~ InoculationLabeled)
RichShootsI2

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedI2 <- contrast(RichShootsI2, adjust = "none")
ShootsPlannedI2

```

**Outlier analysis: below-ground biomass**
```{r}
bio_root = subset(dat, Type == "Root")
  
#interaction model
root_mass1= aov(WeightGrams ~ Inoculation * Water, data = bio_root)
Anova(root_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate

#additive model
root_mass2 = aov(WeightGrams ~ Inoculation + Water, data = bio_root)
Anova(root_mass2, type = "2")

#plot to check residuals
plot(root_mass1)
plot(root_mass2)

# Studentized residuals
MASS::studres(root_mass1) |>
  abs() |>
  sort()

#remove the outliers, re-run the analysis
filtered_root = bio_root[-c(219), ]

```

**Outlier analysis: above-ground biomass**
```{r}

bio_shoot = subset(dat, Type == "Shoot")

#interaction model
shoot_mass1 = aov(WeightGrams ~ Inoculation * Water, data = bio_shoot)
Anova(shoot_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate

#additive model
shoot_mass2 = aov(WeightGrams ~ Inoculation + Water, data = bio_shoot)
Anova(shoot_mass2, type = "2")

#plot to check residuals
plot(shoot_mass1)
plot(shoot_mass2)

# Studentized residuals
MASS::studres(shoot_mass1) |>
  abs() |>
  sort()

#remove the outliers, re-run the analysis
filtered_shoot = bio_shoot[-c(384, 439), ]
```

**ANOVA: below-ground biomass**
```{r}
  
#interaction model
root_mass1= aov(WeightGrams ~ InoculationLabeled * WaterLabeled, data = filtered_root)
Anova(root_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(root_mass1, main = "Root")
plot(allEffects(root_mass1))

#additive model
root_mass2 = aov(WeightGrams ~ InoculationLabeled + WaterLabeled, data = filtered_root)
Anova(root_mass2, type = "2")
plot(root_mass2, main = "Root")
plot(allEffects(root_mass2))
```

**Post-hoc test: below-ground biomass**
```{r}
## Type III ANOVA ##

#apply the emmeans() command to the ANOVA object.
RootsW1<- emmeans(root_mass1, ~ WaterLabeled)
RootsW1

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedW1 <- contrast(RootsW1, adjust = "none")
RootsPlannedW1

#apply the emmeans() command to the ANOVA object.
RootsI1<- emmeans(root_mass1, ~ InoculationLabeled)
RootsI1

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedI1 <- contrast(RootsI1, adjust = "none")
RootsPlannedI1

## Type II ANOVA ##

#apply the emmeans() command to the ANOVA object.
RootsW2<- emmeans(root_mass2, ~ WaterLabeled)
RootsW2

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedW2 <- contrast(RootsW2, adjust = "none")
RootsPlannedW2

#apply the emmeans() command to the ANOVA object.
RootsI2<- emmeans(root_mass2, ~ InoculationLabeled)
RootsI2

#obtain the planned 95% confidence intervals for a pairwise comparison
RootsPlannedI2 <- contrast(RootsI2, adjust = "none")
RootsPlannedI2
```

**ANOVA: above-ground biomass**
```{r}

#interaction model
shoot_mass1 = aov(WeightGrams ~ InoculationLabeled * WaterLabeled, data = filtered_shoot)
Anova(shoot_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(shoot_mass1, main = "Shoot")
plot(allEffects(shoot_mass1))

#additive model
shoot_mass2 = aov(WeightGrams ~ InoculationLabeled + WaterLabeled, data = filtered_shoot)
Anova(shoot_mass2, type = "2")
plot(shoot_mass2, main = "Shoot")
plot(allEffects(shoot_mass2))
```

**Post-hoc test: above-ground biomass**
```{r}
## Type III ANOVA ##

#apply the emmeans() command to the ANOVA object.
ShootsW1<- emmeans(shoot_mass1, ~ WaterLabeled)
ShootsW1

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedW1 <- contrast(ShootsW1, adjust = "none")
ShootsPlannedW1

#apply the emmeans() command to the ANOVA object.
ShootsI1<- emmeans(shoot_mass1, ~ InoculationLabeled)
ShootsI1

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedI1 <- contrast(ShootsI1, adjust = "none")
ShootsPlannedI1

## Type II ANOVA ##

#apply the emmeans() command to the ANOVA object.
ShootsW2<- emmeans(shoot_mass2, ~ WaterLabeled)
ShootsW2

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedW2 <- contrast(ShootsW2, adjust = "none")
ShootsPlannedW2

#apply the emmeans() command to the ANOVA object.
ShootsI2<- emmeans(shoot_mass2, ~ InoculationLabeled)
ShootsI2

#obtain the planned 95% confidence intervals for a pairwise comparison
ShootsPlannedI2 <- contrast(ShootsI2, adjust = "none")
ShootsPlannedI2
```
**PERMANOVA**
```{r, fig.show='hide', warning=FALSE, message=FALSE, include=FALSE}
#load libraries / data
library(labdsv)
library(vegan)
library(phyloseq)
library(ggplot2)

ps.pips <- readRDS("./combined_treatment_data/ps.pips2.rds")
```

```{r}
# run inoculation, then water, then main effects in combined model, then interaction

#bray curtis distance matrix
ps.pips.bray <- phyloseq::distance(ps.pips, method = "bray")

#make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps.pips))
```

**Inoculation effect** 
```{r}
# ---------------------------
#run PERMANOVA for inoculation as the independent variable
adonis_inoculation = adonis2(ps.pips.bray ~ Inoculation, permutations = 9999, data = sampledf)
print(adonis_inoculation)

# Homogeneity of dispersion test
beta.Inoc <- betadisper(ps.pips.bray, sampledf$Inoculation)
permutest(beta.Inoc)
```

**Water effect** 
```{r}
# ---------------------------

#run PERMANOVA water
adonis_water = adonis2(ps.pips.bray ~ Water, permutations = 9999, data = sampledf)
print(adonis_water)

# Homogeneity of dispersion test
beta.Wat <- betadisper(ps.pips.bray, sampledf$Water)
permutest(beta.Wat)
```

**Inoculation + Water effect** 
```{r}
# ---------------------------

# with main effects, no interacton
adonis_combo = adonis2(ps.pips.bray ~ Inoculation + Water, permutations = 9999, data = sampledf, method = "bray")
print(adonis_combo)

```

**Inoculation x Water effect** 
This model shows all levels are significant, but inoculation is most important at R2 = 21%

```{r}
# with interacton
# all levels significant, but inoculation most important R2=21%
adonis_combo = adonis2(ps.pips.bray ~ Inoculation * Water, permutations = 9999, data = sampledf, method = "bray")
print(adonis_combo)


```

**subset the phyloseq objects to get individual objects for each level of inocula.**  
Make sure to subset both the samples but also prune the taxa!
*Note* there are ~200 taxa that would disappear from the taxonomy table if you do a prune taxa for the global dataset with taxa removed if not present in at least one sample: such as `sub.global<-prune_taxa(taxa_sums(ps.pips) > 0, ps.pips)`  
I did this, but it could be reversed just as well.

```{r, warning=FALSE, message=FALSE}
##########
########## Subset dataframe and run permanova for each inoculum testing for water effect

# remove taxa that are not in at least 1 sample globally?
sub.global<-prune_taxa(taxa_sums(ps.pips) > 0, ps.pips)

#subset phyloseq object to make 5 dataframes representing each inoculation treatment
sub1<-subset_samples(sub.global, Inoculation=="1") # only inoculum 1
sub.1<-prune_taxa(taxa_sums(sub1) > 0, sub1) # remove the taxa that are not in at least 1 sample of inoc. 1

sub2<-subset_samples(sub.global, Inoculation=="2") # only inoculum 2
sub.2<-prune_taxa(taxa_sums(sub2) > 0, sub2)

sub3<-subset_samples(sub.global, Inoculation=="3") # only inoculum 3
sub.3<-prune_taxa(taxa_sums(sub3) > 0, sub3)

sub4<-subset_samples(sub.global, Inoculation=="4") # only inoculum 4
sub.4<-prune_taxa(taxa_sums(sub4) > 0, sub4)

sub5<-subset_samples(sub.global, Inoculation=="5") # only inoculum 5
sub.5<-prune_taxa(taxa_sums(sub5) > 0, sub5)

sub6<-subset_samples(sub.global, Inoculation=="6") # only inoculum 6
sub.6<-prune_taxa(taxa_sums(sub6) > 0, sub6)

# now make 5 bray-curtis distance
ps.pips.bray1 <- phyloseq::distance(sub.1, method = "bray")
ps.pips.bray2 <- phyloseq::distance(sub.2, method = "bray")
ps.pips.bray3 <- phyloseq::distance(sub.3, method = "bray")
ps.pips.bray4 <- phyloseq::distance(sub.4, method = "bray")
ps.pips.bray5 <- phyloseq::distance(sub.5, method = "bray")
ps.pips.bray6 <- phyloseq::distance(sub.6, method = "bray")

#make a data frame from the sample_data
sampledf1 <- data.frame(sample_data(sub.1))
sampledf2 <- data.frame(sample_data(sub.2))
sampledf3 <- data.frame(sample_data(sub.3))
sampledf4 <- data.frame(sample_data(sub.4))
sampledf5 <- data.frame(sample_data(sub.5))
sampledf6 <- data.frame(sample_data(sub.6))
```

## Inocula PERMANOVA  
Run the PERMANOVA and betadispersion test for each inocula, testing the effect of water
  
### Inoculum 1
```{r}
### Inoculum 1
adonis_inoc1 = adonis2(ps.pips.bray1 ~ Water, permutations = 9999, data = sampledf1)
print(adonis_inoc1)

beta.1 <- betadisper(ps.pips.bray1, sampledf1$Water)
permutest(beta.1)
```

### Inoculum 2
```{r}
### Inoculum 2
adonis_inoc2 = adonis2(ps.pips.bray2 ~ Water, permutations = 9999, data = sampledf2)
print(adonis_inoc2)

beta.2 <- betadisper(ps.pips.bray2, sampledf2$Water)
permutest(beta.2)
```

### Inoculum 3
```{r}
### Inoculum 3
adonis_inoc3 = adonis2(ps.pips.bray3 ~ Water, permutations = 9999, data = sampledf3)
print(adonis_inoc3)

beta.3 <- betadisper(ps.pips.bray3, sampledf3$Water)
permutest(beta.3)
```

### Inoculum 4
```{r}
### Inoculum 4
adonis_inoc4 = adonis2(ps.pips.bray4 ~ Water, permutations = 9999, data = sampledf4)
print(adonis_inoc4)

beta.4 <- betadisper(ps.pips.bray4, sampledf4$Water)
permutest(beta.4)
```

### Inoculum 5
```{r}
### Inoculum 5
adonis_inoc5 = adonis2(ps.pips.bray5 ~ Water, permutations = 9999, data = sampledf5)
print(adonis_inoc5)

beta.5 <- betadisper(ps.pips.bray5, sampledf5$Water)
permutest(beta.5)

#########################

```

### Inoculum 6
```{r}
### Inoculum 6
adonis_inoc6 = adonis2(ps.pips.bray6 ~ Water, permutations = 9999, data = sampledf6)
print(adonis_inoc6)

beta.6 <- betadisper(ps.pips.bray6, sampledf6$Water)
permutest(beta.6)
```