---
title: "PERMANOVA"
author: "Hynson Lab"
date: "11/11/2021"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---
  
```{r setup chunk, setup, warning=FALSE, message=FALSE, fig.align='center', include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**PERMANOVA**
```{r, fig.show='hide', warning=FALSE, message=FALSE, include=FALSE}
#load libraries / data
library(labdsv)
library(vegan)
library(phyloseq)
library(ggplot2)

ps.pips <- readRDS("./intermediate_data/ps.pips.rds")
```

```{r}
# run inoculation, then water, then main effects in combined model, then interaction

#bray curtis distance matrix
ps.pips.bray <- phyloseq::distance(ps.pips, method = "bray")

#make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps.pips))
```

**Inoculation effect** 
```{r}
# ---------------------------
#run PERMANOVA for inoculation as the independent variable
adonis_inoculation = adonis2(ps.pips.bray ~ Inoculation, permutations = 9999, data = sampledf)
print(adonis_inoculation)

# Homogeneity of dispersion test
beta.Inoc <- betadisper(ps.pips.bray, sampledf$Inoculation)
permutest(beta.Inoc)
```

**Water effect** 
```{r}
# ---------------------------

#run PERMANOVA water
adonis_water = adonis2(ps.pips.bray ~ Water, permutations = 9999, data = sampledf)
print(adonis_water)

# Homogeneity of dispersion test
beta.Wat <- betadisper(ps.pips.bray, sampledf$Water)
permutest(beta.Wat)
```

**Inoculation + Water effect** 
```{r}
# ---------------------------

# with main effects, no interacton
adonis_combo = adonis2(ps.pips.bray ~ Inoculation + Water, permutations = 9999, data = sampledf, method = "bray")
print(adonis_combo)

```

**Inoculation x Water effect** 
This model shows all levels are significant, but inoculation is most important at R2 = 21%

```{r}
# with interacton
# all levels significant, but inoculation most important R2=21%
adonis_combo = adonis2(ps.pips.bray ~ Inoculation * Water, permutations = 9999, data = sampledf, method = "bray")
print(adonis_combo)


```

**subset the phyloseq objects to get individual objects for each level of inocula.**  
Make sure to subset both the samples but also prune the taxa!
*Note* there are ~200 taxa that would disappear from the taxonomy table if you do a prune taxa for the global dataset with taxa removed if not present in at least one sample: such as `sub.global<-prune_taxa(taxa_sums(ps.pips) > 0, ps.pips)`  
I did this, but it could be reversed just as well.

```{r, warning=FALSE, message=FALSE}
##########
########## Subset dataframe and run permanova for each inoculum testing for water effect

# remove taxa that are not in at least 1 sample globally?
sub.global<-prune_taxa(taxa_sums(ps.pips) > 0, ps.pips)

#subset phyloseq object to make 5 dataframes representing each inoculation treatment
sub1<-subset_samples(sub.global, Inoculation=="1") # only inoculum 1
sub.1<-prune_taxa(taxa_sums(sub1) > 0, sub1) # remove the taxa that are not in at least 1 sample of inoc. 1

sub2<-subset_samples(sub.global, Inoculation=="2") # only inoculum 2
sub.2<-prune_taxa(taxa_sums(sub1) > 0, sub2)

sub3<-subset_samples(sub.global, Inoculation=="3") # only inoculum 3
sub.3<-prune_taxa(taxa_sums(sub3) > 0, sub3)

sub4<-subset_samples(sub.global, Inoculation=="4") # only inoculum 4
sub.4<-prune_taxa(taxa_sums(sub4) > 0, sub4)

sub5<-subset_samples(sub.global, Inoculation=="5") # only inoculum 5
sub.5<-prune_taxa(taxa_sums(sub1) > 0, sub5)


# now make 5 bray-curtis distance
ps.pips.bray1 <- phyloseq::distance(sub.1, method = "bray")
ps.pips.bray2 <- phyloseq::distance(sub.2, method = "bray")
ps.pips.bray3 <- phyloseq::distance(sub.3, method = "bray")
ps.pips.bray4 <- phyloseq::distance(sub.4, method = "bray")
ps.pips.bray5 <- phyloseq::distance(sub.5, method = "bray")

#make a data frame from the sample_data
sampledf1 <- data.frame(sample_data(sub.1))
sampledf2 <- data.frame(sample_data(sub.2))
sampledf3 <- data.frame(sample_data(sub.3))
sampledf4 <- data.frame(sample_data(sub.4))
sampledf5 <- data.frame(sample_data(sub.5))
```

## Inocula PERMANOVA  
Run the PERMANOVA and betadispersion test for each inocula, testing the effect of water
  
### Inoculum 1
```{r}
### Inoculum 1
adonis_inoc1 = adonis2(ps.pips.bray1 ~ Water, permutations = 9999, data = sampledf1)
print(adonis_inoc1)

beta.1 <- betadisper(ps.pips.bray1, sampledf1$Water)
permutest(beta.1)
```

### Inoculum 2
```{r}
### Inoculum 2
adonis_inoc2 = adonis2(ps.pips.bray2 ~ Water, permutations = 9999, data = sampledf2)
print(adonis_inoc2)

beta.2 <- betadisper(ps.pips.bray2, sampledf2$Water)
permutest(beta.2)
```

### Inoculum 3
```{r}
### Inoculum 3
adonis_inoc3 = adonis2(ps.pips.bray3 ~ Water, permutations = 9999, data = sampledf3)
print(adonis_inoc3)

beta.3 <- betadisper(ps.pips.bray3, sampledf3$Water)
permutest(beta.3)
```

### Inoculum 4
```{r}
### Inoculum 4
adonis_inoc4 = adonis2(ps.pips.bray4 ~ Water, permutations = 9999, data = sampledf4)
print(adonis_inoc4)

beta.4 <- betadisper(ps.pips.bray4, sampledf4$Water)
permutest(beta.4)
```

### Inoculum 5
```{r}
### Inoculum 5
adonis_inoc5 = adonis2(ps.pips.bray5 ~ Water, permutations = 9999, data = sampledf5)
print(adonis_inoc5)

beta.5 <- betadisper(ps.pips.bray5, sampledf5$Water)
permutest(beta.5)

#########################

```