---
title: "Strawberry Analysis"
author: "Francisca Rodriguez"
date: "9/4/2021"
output: html_document
---

``````{r global options, results="hide", warning=FALSE, message=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')
 
# Load in packages
if (!require("pacman")) install.packages("pacman"); library(pacman) # for rapid install if not in library
if(!requireNamespace("BiocManager", quietly = TRUE))
	install.packages("BiocManager")	#Load packages from Bioconductor


#Packages
pacman::p_load('FedData', 'latticeExtra', 'scales', 'viridis', 'ggmap', 'phyloseq', 'raster', 'rgdal', 'RColorBrewer', 'ggplot2', 'vegan', 'ggplot2', 'phyloseq', 'bipartite', 'plotrix', 'viridis', 'lattice', 'fossil', 'plyr', 'devtools', 'ggpubr', 'gridExtra', 'cowplot', 'rgr', 'ggplot2', 'cowplot', 'tidyverse', 'dada2', 'purrr', 'Biostrings', 'gapminder', 'tibble', 'ggridges', 'hrbrthemes', 'decontam', 'iNEXT')
```


**Data input and Phyloseq**  
Input the data files from dada2 sequence table (chimeras removed) and the dada2 taxonomy and Miseq map for metadata. Use these and make a complete csv that has non-run or unnecessary samples removed and label items/levels/treatments to help in down stream processing. 

Subsequently, build the phyloseq data tables and generate ASV names with taxonomy, export the object and R.data

```{r, initial setup}
##INITIAL SETUP################################

#Load seqtable.nochim and taxa from dada2 files from Kacie
seqtable.nochim <- readRDS('data/raw_data/dada2_seqtable.RDS')
taxa <- readRDS('data/raw_data/RDP_Taxonomy_from_dada2.RDS')
strawb_miseq_map <- read.csv("data/raw_data/Strawberry_Miseq_Map_forQIIME.csv", fileEncoding="UTF-8-BOM")

# add empty column
strawb_miseq_map <- add_column(strawb_miseq_map, SampleType = NA)

# get rid of unneeded cols
strawb_miseq_map <- select(strawb_miseq_map, 
                           SampleID, 
                           TimePoint, 
                           Type, 
                           PlantID, 
                           WeightGrams, 
                           SampleNotes, 
                           Water, 
                           Inoculation, 
                           Replicate, 
                           PlantNotes, 
                           SampleType)

# format Water and replicate as factor
strawb_miseq_map$Water<-as.factor(strawb_miseq_map$Water)
strawb_miseq_map$Replicate<-as.factor(strawb_miseq_map$Replicate)

# label sample types
strawb_miseq_map$SampleType[strawb_miseq_map$TimePoint == "T0"] <- "post_inoc_pre_stress"
strawb_miseq_map$SampleType[strawb_miseq_map$TimePoint == "T1"] <- "post_inoc_post_stress"
strawb_miseq_map$SampleType[strawb_miseq_map$SampleNotes == "Empty Extraction Tube Control"] <- "DNA_PCR_neg"
strawb_miseq_map$SampleType[strawb_miseq_map$SampleNotes == "Empty Extraction Tube Control"] <- "DNA_PCR_neg"
strawb_miseq_map$SampleType[grep("Inoc", strawb_miseq_map$SampleNotes)] <- "slurry"
#strawb_miseq_map$SampleType[strawb_miseq_map$Inoculation == "6"] <- "i6"
strawb_miseq_map$SampleType[grep("Inoculum 6", strawb_miseq_map$SampleNotes)] <- "di_water"
strawb_miseq_map$SampleType[grep("pre-inoc", strawb_miseq_map$SampleNotes)] <- "pre_inoc"

# add empty column for control
strawb_miseq_map <- add_column(strawb_miseq_map, Sample_or_Control = "True Sample")

#label controls (i6, DI water, DNA/PCR negative, pre-inoculated plants)
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$Inoculation == "6"] <- "Control"
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$SampleType == "di_water"] <- "Control"
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$SampleType == "DNA_PCR_neg"] <- "Control"
strawb_miseq_map$Sample_or_Control[strawb_miseq_map$SampleType == "pre_inoc"] <- "Control"

# remove unwanted samples
strawb_miseq_map <- strawb_miseq_map[grep("CSIA", strawb_miseq_map$SampleNotes, invert = TRUE),]
strawb_miseq_map <- strawb_miseq_map[strawb_miseq_map$Type != "Palmyra",]

#save new map as a .csv file
write.csv(strawb_miseq_map, "data/raw_data/strawb_miseq_map.csv")
```


```{r, phyloseq}
#############################################################################
#################### PHYLOSEQ ###############################################
#############################################################################

# Hand off to Phyloseq ####
otu <- otu_table(seqtable.nochim,taxa_are_rows = FALSE)
tax <- tax_table(taxa)
met <- sample_data(strawb_miseq_map)
row.names(met) <- strawb_miseq_map$SampleID

ps <- phyloseq(otu,met,tax)

# generate the ASV0001, ASV0002 names
pretty_names <- paste("ASV", sprintf('%0.4d', 1:length(taxa_names(ps))), sep = "")

# reassign the taxa_names from the sequences to the generated ASV names
taxa_names(ps) <- pretty_names


inspect.richness<-plot_richness(ps, x="Inoculation", color="Inoculation", measures="Observed")

inspect.richness
dev.copy(pdf, "figures/inspect.richness.pdf", height=4, width=6)
dev.off() 
```

**Pre-decontamination**  
Before we decontaminate the samples we will assess otu tables for different samples and explore these with Venn diagrams. Load in the phyloseq object and subset by sample type. From here we will assess the negative and positive controls  

```{r, pre decontam Venn, eval=FALSE}
#############################################################################
############  PRE-DECONTAM: VENN DIAGRAMS  ##################################
#############################################################################
library(ggVennDiagram)

# pre inoc table
pre_inoc_ps <- subset_samples(ps, sample_data(ps)$SampleType=="pre_inoc")
pre_inoc_asv <- as.data.frame(otu_table(pre_inoc_ps))

#slurry table
slurry_ps <- subset_samples(ps, sample_data(ps)$SampleType=="slurry")
slurry_asv <- as.data.frame(otu_table(slurry_ps))

# #DNA_PCR_neg table
DNA_PCR_neg_ps <- subset_samples(ps, sample_data(ps)$SampleType=="DNA_PCR_neg")
DNA_PCR_neg_asv <- as.data.frame(otu_table(DNA_PCR_neg_ps))

#di_water table
di_water_ps <- subset_samples(ps, sample_data(ps)$SampleType=="di_water")
di_water_asv <- as.data.frame(otu_table(di_water_ps))

#i6 table
i6_ps <- subset_samples(ps, sample_data(ps)$SampleType=="i6")
i6_asv <- as.data.frame(otu_table(i6_ps))

#post_inoc_pre_stress table
post_inoc_pre_stress_ps <- subset_samples(ps, sample_data(ps)$SampleType=="post_inoc_pre_stress")
post_inoc_pre_stress_asv <- as.data.frame(otu_table(post_inoc_pre_stress_ps))

##subset post_inoc_pre_stress to count ############################################## 
post_inoc_pre_stress_i1 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "1")
post_inoc_pre_stress_i2 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "2")
post_inoc_pre_stress_i3 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "3")
post_inoc_pre_stress_i4 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "4")
post_inoc_pre_stress_i5 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "5")
post_inoc_pre_stress_i6 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "6")

#post_inoc_post_stress table
post_inoc_post_stress_ps <- subset_samples(ps, sample_data(ps)$SampleType=="post_inoc_post_stress")
post_inoc_post_stress_asv <- as.data.frame(otu_table(post_inoc_post_stress_ps))

##subset post_inoc_post_stress to count###############################################
#inoculation 1 
post_inoc_post_stress_i1_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "1" & sample_data(post_inoc_post_stress_ps)$Water == "1")
post_inoc_post_stress_i1_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "1" & sample_data(post_inoc_post_stress_ps)$Water == "2")
post_inoc_post_stress_i1_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "1" & sample_data(post_inoc_post_stress_ps)$Water == "3")

#inoculation 2
post_inoc_post_stress_i2_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "2" & sample_data(post_inoc_post_stress_ps)$Water == "1")
post_inoc_post_stress_i2_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "2" & sample_data(post_inoc_post_stress_ps)$Water == "2")
post_inoc_post_stress_i2_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "2" & sample_data(post_inoc_post_stress_ps)$Water == "3")

#inoculation 3
post_inoc_post_stress_i3_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "3" & sample_data(post_inoc_post_stress_ps)$Water == "1")
post_inoc_post_stress_i3_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "3" & sample_data(post_inoc_post_stress_ps)$Water == "2")
post_inoc_post_stress_i3_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "3" & sample_data(post_inoc_post_stress_ps)$Water == "3")

#inoculation 4
post_inoc_post_stress_i4_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "4" & sample_data(post_inoc_post_stress_ps)$Water == "1")
post_inoc_post_stress_i4_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "4" & sample_data(post_inoc_post_stress_ps)$Water == "2")
post_inoc_post_stress_i4_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "4" & sample_data(post_inoc_post_stress_ps)$Water == "3")

#inoculation 5
post_inoc_post_stress_i5_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "5" & sample_data(post_inoc_post_stress_ps)$Water == "1")
post_inoc_post_stress_i5_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "5" & sample_data(post_inoc_post_stress_ps)$Water == "2")
post_inoc_post_stress_i5_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "5" & sample_data(post_inoc_post_stress_ps)$Water == "3")

#inoculation 6
post_inoc_post_stress_i6_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "6" & sample_data(post_inoc_post_stress_ps)$Water == "1")
post_inoc_post_stress_i6_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "6" & sample_data(post_inoc_post_stress_ps)$Water == "2")
post_inoc_post_stress_i6_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "6" & sample_data(post_inoc_post_stress_ps)$Water == "3")

#############################################################################
##NEGATIVE CONTROL EXPOLORATION - VENN DIAGRAMS##############################
#############################################################################

#sum asv columns into a list
sum_pre_inoc_asv = colSums(pre_inoc_asv)
sum_slurry_asv = colSums(slurry_asv)
sum_DNA_PCR_neg_asv = colSums(DNA_PCR_neg_asv)
sum_di_water_asv = colSums(di_water_asv)
sum_i6_asv = colSums(i6_asv)
sum_post_inoc_pre_stress_asv = colSums(post_inoc_pre_stress_asv)
sum_post_inoc_post_stress_asv = colSums(post_inoc_post_stress_asv)

#get index of values > 0
index_pre_inoc_asv = which(sum_pre_inoc_asv > 0)
index_slurry_asv = which(sum_slurry_asv > 0)
index_DNA_PCR_neg_asv = which(sum_DNA_PCR_neg_asv > 0)
index_di_water_asv = which(sum_di_water_asv > 0)
index_i6_asv = which(sum_i6_asv > 0)
index_post_inoc_pre_stress_asv = which(sum_post_inoc_pre_stress_asv > 0)
index_post_inoc_post_stress_asv = which(sum_post_inoc_post_stress_asv > 0)

#create a list containing all the indexes from above
total_index <- list(DNA_PCR_neg = index_DNA_PCR_neg_asv, di_water = index_di_water_asv, pre_inoc = index_pre_inoc_asv, slurry = index_slurry_asv, post_inoc_pre_stress = index_post_inoc_pre_stress_asv, post_inoc_post_stress = index_post_inoc_post_stress_asv, i6 = index_i6_asv)

#create venn diagram
ggVennDiagram(total_index, label = "count")


#find highest number of reads for any ASV found in each sample type
max_pre_inoc_asv = max(sum_pre_inoc_asv)
max_slurry_asv = max(sum_slurry_asv)
max_DNA_PCR_neg_asv = max(sum_DNA_PCR_neg_asv)
max_di_water_asv = max(sum_di_water_asv)
max_i6_asv = max(sum_i6_asv)
max_post_inoc_pre_stress_asv = max(sum_post_inoc_pre_stress_asv)
max_post_inoc_post_stress_asv = max(sum_post_inoc_post_stress_asv)

#check total count of ASVs, remove ASVs with one count, re-tally
ASV_table <- as(otu_table(ps), "matrix")
ASV_sums <- colSums(ASV_table)
ASV_index <- which(ASV_sums > 1)

```
  
```{r, pre decontam richness, eval=FALSE}
#############################################################################
####### PRE-DECONTAM: DETERMINE MEAN/STD. E OF SAMPLES ######################
#############################################################################
#Now calculate richness / mean / se / replicate number (n) for each inoculum and treatment after the stress period.  

##subset slurry by inoculation ##############################################
slurry1 <- subset_samples(slurry_ps, sample_data(slurry_ps)$Type == "Inoculum 1")
slurry1_asv <- as.data.frame(otu_table(slurry1))
sum_slurry1 <- colSums(slurry1_asv)
  
slurry2 <- subset_samples(slurry_ps, sample_data(slurry_ps)$Type == "Inoculum 2")
slurry2_asv <- as.data.frame(otu_table(slurry2))
sum_slurry2 <- colSums(slurry2_asv)

slurry3 <- subset_samples(slurry_ps, sample_data(slurry_ps)$Type == "Inoculum 3")
slurry3_asv <- as.data.frame(otu_table(slurry3))
sum_slurry3 <- colSums(slurry3_asv)

slurry4 <- subset_samples(slurry_ps, sample_data(slurry_ps)$Type == "Inoculum 4")
slurry4_asv <- as.data.frame(otu_table(slurry4))
sum_slurry4 <- colSums(slurry4_asv)

slurry5 <- subset_samples(slurry_ps, sample_data(slurry_ps)$Type == "Inoculum 5")
slurry5_asv <- as.data.frame(otu_table(slurry5))
sum_slurry5 <- colSums(slurry5_asv)

##subset post_inoc_post_stress by water treatment ###########################
post_inoc_post_stress1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Water == "1")
post_inoc_post_stress1_asv <- as.data.frame(otu_table(post_inoc_post_stress1))
sum_post_inoc_post_stress1 <- colSums(post_inoc_post_stress1_asv)

post_inoc_post_stress2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Water == "2")
post_inoc_post_stress2_asv <- as.data.frame(otu_table(post_inoc_post_stress2))
sum_post_inoc_post_stress2 <- colSums(post_inoc_post_stress2_asv)

post_inoc_post_stress3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Water == "3")
post_inoc_post_stress3_asv <- as.data.frame(otu_table(post_inoc_post_stress3))
sum_post_inoc_post_stress3 <- colSums(post_inoc_post_stress3_asv)

#find avg. number of ASVs
mean_pre_inoc_asv = mean(sum_pre_inoc_asv)
mean_di_water_asv = mean(sum_di_water_asv)
mean_DNA_PCR_neg_asv = mean(sum_DNA_PCR_neg_asv)
mean_i6_asv = mean(sum_i6_asv)
mean_slurry1_asv = mean(sum_slurry1)
mean_slurry2_asv = mean(sum_slurry2)
mean_slurry3_asv = mean(sum_slurry3)
mean_slurry4_asv = mean(sum_slurry4)
mean_slurry5_asv = mean(sum_slurry5)
mean_post_inoc_pre_stress_asv = mean(sum_post_inoc_pre_stress_asv)
mean_post_inoc_post_stress1_asv = mean(sum_post_inoc_post_stress1)
mean_post_inoc_post_stress2_asv = mean(sum_post_inoc_post_stress2)
mean_post_inoc_post_stress3_asv = mean(sum_post_inoc_post_stress3)

#find std. err. of ASVs
std_err_pre_inoc_asv = sd(sum_pre_inoc_asv) / sqrt(length(sum_pre_inoc_asv))
std_err_di_water_asv = sd(sum_di_water_asv) / sqrt(length(sum_di_water_asv))
std_err_DNA_PCR_neg_asv = sd(sum_DNA_PCR_neg_asv) / sqrt(length(sum_DNA_PCR_neg_asv))
std_err_i6_asv = sd(sum_i6_asv) / sqrt(length(sum_i6_asv))
std_err_slurry1_asv = sd(sum_slurry1) / sqrt(length(sum_slurry1))
std_err_slurry2_asv = sd(sum_slurry2) / sqrt(length(sum_slurry2))
std_err_slurry3_asv = sd(sum_slurry3) / sqrt(length(sum_slurry3))
std_err_slurry4_asv = sd(sum_slurry4) / sqrt(length(sum_slurry4))
std_err_slurry5_asv = sd(sum_slurry5) / sqrt(length(sum_slurry5))
std_err_post_inoc_pre_stress_asv = sd(sum_post_inoc_pre_stress_asv) / sqrt(length(sum_post_inoc_pre_stress_asv))
std_err_post_inoc_post_stress1_asv = sd(sum_post_inoc_post_stress1) / sqrt(length(sum_post_inoc_post_stress1))
std_err_post_inoc_post_stress2_asv = sd(sum_post_inoc_post_stress2) / sqrt(length(sum_post_inoc_post_stress2))
std_err_post_inoc_post_stress3_asv = sd(sum_post_inoc_post_stress3) / sqrt(length(sum_post_inoc_post_stress3))

#create a dataframe to export control data ##################################
# these are all samples from DI water, negaive controls and the slurries

ctrl_means <- c(mean_pre_inoc_asv, 
           mean_di_water_asv, 
           mean_DNA_PCR_neg_asv, 
           mean_i6_asv, 
           mean_slurry1_asv, 
           mean_slurry2_asv, 
           mean_slurry3_asv, 
           mean_slurry4_asv, 
           mean_slurry5_asv)

ctrl_std_err <- c(std_err_pre_inoc_asv, 
             std_err_di_water_asv, 
             std_err_DNA_PCR_neg_asv, 
             std_err_i6_asv, 
             std_err_slurry1_asv, 
             std_err_slurry2_asv, 
             std_err_slurry3_asv, 
             std_err_slurry4_asv, 
             std_err_slurry5_asv)

ctrl_labels <- c("pre_inoc", 
            "di_water", 
            "DNA_PCR_neg", 
            "i6", 
            "slurry1", 
            "slurry2", 
            "slurry3", 
            "slurry4", 
            "slurry5")
            
ctrl.mean.SE.ASVs <- rbind(ctrl_labels, ctrl_means, ctrl_std_err)

#export as .csv
ctrl.mean.SE.ASVs <-  write.csv(ctrl.mean.SE.ASVs, "output/ctrl.mean.SE.ASVs.csv")

## repeat for experimental data #############################################
## these are samples after the water stress component

exp_means <- c(mean_post_inoc_pre_stress_asv, 
                  mean_post_inoc_post_stress1_asv, 
                  mean_post_inoc_post_stress2_asv, 
                  mean_post_inoc_post_stress3_asv)

exp_std_err <- c(std_err_post_inoc_pre_stress_asv, 
                  std_err_post_inoc_post_stress1_asv, 
                  std_err_post_inoc_post_stress2_asv, 
                  std_err_post_inoc_post_stress3_asv)

exp_labels <- c("post_inoc_pre_stress",
                  "post_inoc_post_stress1", 
                  "post_inoc_post_stress2", 
                  "post_inoc_post_stress3")
            
exp.mean.SE.ASVs <- rbind(exp_labels, exp_means, exp_std_err)

#export as .csv
exp.mean.SE.ASVs <-  write.csv(exp.mean.SE.ASVs, "output/exp.mean.SE.ASVs.csv")

```

```{r, negative controls, eval=FALSE}
#############################################################################
###### PRE-DECONTAM: CHECK SPREAD OF NEGATIVE CONTROL ASVS  #################
#############################################################################

##DNA PCR NEG####
DNA_PCR_neg_ps <- subset_samples(ps, sample_data(ps)$SampleType=="DNA_PCR_neg") #DNA_PCR_neg table
DNA_PCR_neg_ps <- prune_taxa(taxa_sums(DNA_PCR_neg_ps) > 1, DNA_PCR_neg_ps) #get rid of zeroes
DNA_PCR_neg_names <- paste("DNA_PCR_neg_", sprintf('%0.4d', 1:length(sample_names(DNA_PCR_neg_ps))), sep = "") # generate the row names 
sample_names(DNA_PCR_neg_ps) <- DNA_PCR_neg_names # reassign the names from the sequences to the generated sample names
DNA_PCR_neg_asv <- as.data.frame(otu_table(DNA_PCR_neg_ps)) #create asv table

##DI WATER#############
di_water_ps <- subset_samples(ps, sample_data(ps)$SampleType=="di_water")
di_water_ps <- prune_taxa(taxa_sums(di_water_ps) > 1, di_water_ps)
di_water_names <- paste("di_water_", sprintf('%0.4d', 1:length(sample_names(di_water_ps))), sep = "") 
sample_names(di_water_ps) <- di_water_names
di_water_asv <- as.data.frame(otu_table(di_water_ps))

##INOCULATION 6 - H2O CONTROL#############
i6_ps <- subset_samples(ps, sample_data(ps)$SampleType=="i6")
i6_ps <- prune_taxa(taxa_sums(i6_ps) > 1, i6_ps)
i6_names <- paste("i6_", sprintf('%0.4d', 1:length(sample_names(i6_ps))), sep = "") 
sample_names(i6_ps) <- i6_names
i6_asv <- as.data.frame(otu_table(i6_ps))

##PRE INOC########
pre_inoc_ps <- subset_samples(ps, sample_data(ps)$SampleType=="pre_inoc")
pre_inoc_ps <- prune_taxa(taxa_sums(pre_inoc_ps) > 1, pre_inoc_ps)
pre_inoc_names <- paste("pre_inoc_", sprintf('%0.4d', 1:length(sample_names(pre_inoc_ps))), sep = "") 
sample_names(pre_inoc_ps) <- pre_inoc_names
pre_inoc_asv <- as.data.frame(otu_table(pre_inoc_ps))

##SLURRY#########
slurry_ps <- subset_samples(ps, sample_data(ps)$SampleType=="slurry") #slurry table
slurry_names <- paste("slurry_", sprintf('%0.4d', 1:length(sample_names(slurry_ps))), sep = "")
sample_names(slurry_ps) <- slurry_names 
slurry_asv <- as.data.frame(otu_table(slurry_ps))

##POST INOC PRE STRESS#####
post_inoc_pre_stress_ps <- subset_samples(ps, sample_data(ps)$SampleType=="post_inoc_pre_stress")
post_inoc_pre_stress_names <- paste("post_inoc_pre_stress_", sprintf('%0.4d', 1:length(sample_names(post_inoc_pre_stress_ps))), sep = "")
sample_names(post_inoc_pre_stress_ps) <- post_inoc_pre_stress_names
post_inoc_pre_stress_asv <- as.data.frame(otu_table(post_inoc_pre_stress_ps))

##POST INOC POST STRESS####
post_inoc_post_stress_ps <- subset_samples(ps, sample_data(ps)$SampleType=="post_inoc_post_stress")
post_inoc_post_stress_names <- paste("post_inoc_post_stress_", sprintf('%0.4d', 1:length(sample_names(post_inoc_post_stress_ps))), sep = "")
sample_names(post_inoc_post_stress_ps) <- post_inoc_post_stress_names
post_inoc_post_stress_asv <- as.data.frame(otu_table(post_inoc_post_stress_ps)) 

##COMBINE ASV TABLES########

#create control ASV table
control_asvs <- dplyr::bind_rows(DNA_PCR_neg_asv, di_water_asv, i6_asv, pre_inoc_asv)

#obtain a list of all the columns in that data frame
asvs <- colnames(control_asvs)

#filter for ASVs present in controls
edited_slurry <- slurry_asv[,asvs]
edited_postinocprestress <- post_inoc_pre_stress_asv[,asvs]
edited_postinocpoststress <- post_inoc_post_stress_asv[,asvs]

#combine data frames
comboASVs <- dplyr::bind_rows(control_asvs, edited_slurry, edited_postinocprestress, edited_postinocpoststress)

#export as .csv
controlASVs <-  write.csv(comboASVs, "output/controlASVs.csv")

##DETERMINE TOTAL READS FROM TREATMENT ####
totalTreatmentReads <- rowSums(post_inoc_post_stress_asv)
treatmentASVs <- write.csv(totalTreatmentReads, "output/treatmentASVs.csv")

```

**Decontaminate controls**  
We will use the prevalence method for removing contaminants in DADA2.  

```{r, decontam controls}
#############################################################################
################ DECONTAMINATE CONTROLS #####################################
#############################################################################

#remove ASVs with only one read 
ps <- prune_taxa(taxa_sums(ps) > 1, ps)

# 1964 taxa
ps

#look at the library sizes of the controls and samples
df <- as.data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_or_Control)) + geom_point()

#label controls (extraction negatives) and clean
sample_data(ps)$is.neg <- sample_data(ps)$Sample_or_Control == "Control"
contams <- isContaminant(ps, method="prevalence", neg="is.neg")
table(contams$contaminant)

#Remove contaminants
ps.noncontam <- prune_taxa(!contams$contaminant, ps)
ps.noncontam
```

```{r, asv post decontam, eval=FALSE}
#############################################################################
### POST-DECONTAM: DATA BY PRE/POST STRESS AND INOC X WATER TREATMENT   #####
#############################################################################
#Re-visit ASVs now that decontaminants have been removed.
  


##POST INOC PRE STRESS ######################################################

#post_inoc_pre_stress table 
post_inoc_pre_stress_ps <- subset_samples(ps.noncontam, sample_data(ps.noncontam)$SampleType=="post_inoc_pre_stress")

##label post_inoc_pre_stress by inoculation
post_inoc_pre_stress_i1 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "1")
  post_inoc_pre_stress_i1 <- as.data.frame(otu_table(post_inoc_pre_stress_i1))
  
post_inoc_pre_stress_i2 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "2")
  post_inoc_pre_stress_i2 <- as.data.frame(otu_table(post_inoc_pre_stress_i2))
  
post_inoc_pre_stress_i3 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "3")
  post_inoc_pre_stress_i3 <- as.data.frame(otu_table(post_inoc_pre_stress_i3))
  
post_inoc_pre_stress_i4 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "4")
  post_inoc_pre_stress_i4 <- as.data.frame(otu_table(post_inoc_pre_stress_i4))
  
post_inoc_pre_stress_i5 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "5")
  post_inoc_pre_stress_i5 <- as.data.frame(otu_table(post_inoc_pre_stress_i5))

post_inoc_pre_stress_i6 <- subset_samples(post_inoc_pre_stress_ps, sample_data(post_inoc_pre_stress_ps)$Inoculation == "6")
  post_inoc_pre_stress_i6 <- as.data.frame(otu_table(post_inoc_pre_stress_i6))
  
  
##POST INOC POST STRESS#####################################################

#post_inoc_post_stress table 
post_inoc_post_stress_ps <- subset_samples(ps.noncontam, sample_data(ps.noncontam)$SampleType=="post_inoc_post_stress")

##label post_inoc_post_stress by inoculation x water treatment

#inoculation 1 
post_inoc_post_stress_i1_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "1" & sample_data(post_inoc_post_stress_ps)$Water == "1")
  post_inoc_post_stress_i1_w1 <- as.data.frame(otu_table(post_inoc_post_stress_i1_w1))

post_inoc_post_stress_i1_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "1" & sample_data(post_inoc_post_stress_ps)$Water == "2")
  post_inoc_post_stress_i1_w2 <- as.data.frame(otu_table(post_inoc_post_stress_i1_w2))

post_inoc_post_stress_i1_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "1" & sample_data(post_inoc_post_stress_ps)$Water == "3")
  post_inoc_post_stress_i1_w3 <- as.data.frame(otu_table(post_inoc_post_stress_i1_w3))

#inoculation 2
post_inoc_post_stress_i2_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "2" & sample_data(post_inoc_post_stress_ps)$Water == "1")
  post_inoc_post_stress_i2_w1 <- as.data.frame(otu_table(post_inoc_post_stress_i2_w1))

post_inoc_post_stress_i2_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "2" & sample_data(post_inoc_post_stress_ps)$Water == "2")
  post_inoc_post_stress_i2_w2 <- as.data.frame(otu_table(post_inoc_post_stress_i2_w2))

post_inoc_post_stress_i2_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "2" & sample_data(post_inoc_post_stress_ps)$Water == "3")
  post_inoc_post_stress_i2_w3 <- as.data.frame(otu_table(post_inoc_post_stress_i2_w3))

#inoculation 3
post_inoc_post_stress_i3_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "3" & sample_data(post_inoc_post_stress_ps)$Water == "1")
  post_inoc_post_stress_i3_w1 <- as.data.frame(otu_table(post_inoc_post_stress_i3_w1))

post_inoc_post_stress_i3_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "3" & sample_data(post_inoc_post_stress_ps)$Water == "2")
  post_inoc_post_stress_i3_w2 <- as.data.frame(otu_table(post_inoc_post_stress_i3_w2))

post_inoc_post_stress_i3_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "3" & sample_data(post_inoc_post_stress_ps)$Water == "3")
  post_inoc_post_stress_i3_w3 <- as.data.frame(otu_table(post_inoc_post_stress_i3_w3))

#inoculation 4
post_inoc_post_stress_i4_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "4" & sample_data(post_inoc_post_stress_ps)$Water == "1")
  post_inoc_post_stress_i4_w1 <- as.data.frame(otu_table(post_inoc_post_stress_i4_w1))

post_inoc_post_stress_i4_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "4" & sample_data(post_inoc_post_stress_ps)$Water == "2")
  post_inoc_post_stress_i4_w2 <- as.data.frame(otu_table(post_inoc_post_stress_i4_w2))

post_inoc_post_stress_i4_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "4" & sample_data(post_inoc_post_stress_ps)$Water == "3")
  post_inoc_post_stress_i4_w3 <- as.data.frame(otu_table(post_inoc_post_stress_i4_w3))

#inoculation 5
post_inoc_post_stress_i5_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "5" & sample_data(post_inoc_post_stress_ps)$Water == "1")
  post_inoc_post_stress_i5_w1 <- as.data.frame(otu_table(post_inoc_post_stress_i5_w1))

post_inoc_post_stress_i5_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "5" & sample_data(post_inoc_post_stress_ps)$Water == "2")
  post_inoc_post_stress_i5_w2 <- as.data.frame(otu_table(post_inoc_post_stress_i5_w2))

post_inoc_post_stress_i5_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "5" & sample_data(post_inoc_post_stress_ps)$Water == "3")
  post_inoc_post_stress_i5_w3 <- as.data.frame(otu_table(post_inoc_post_stress_i5_w3))

#inoculation 6
post_inoc_post_stress_i6_w1 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "6" & sample_data(post_inoc_post_stress_ps)$Water == "1")
  post_inoc_post_stress_i6_w1 <- as.data.frame(otu_table(post_inoc_post_stress_i6_w1))

post_inoc_post_stress_i6_w2 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "6" & sample_data(post_inoc_post_stress_ps)$Water == "2")
  post_inoc_post_stress_i6_w2 <- as.data.frame(otu_table(post_inoc_post_stress_i6_w2))

post_inoc_post_stress_i6_w3 <- subset_samples(post_inoc_post_stress_ps, sample_data(post_inoc_post_stress_ps)$Inoculation == "6" & sample_data(post_inoc_post_stress_ps)$Water == "3")
  post_inoc_post_stress_i6_w3 <- as.data.frame(otu_table(post_inoc_post_stress_i6_w3))

```
```{r, species curves beta diversity, eval=FALSE}
#############################################################################
#####  POST-DECONTAM: SPECIES ACCUMULATION CURVES (BETA DIVERSITY)  #########
#############################################################################
# Species Curves**  
# Here we will be assessing the accumulation curves for samples before and after stress.  


par(mfrow=c(4,5), mar=c(1,1,1,1))

##POST INOC PRE STRESS #####################################################
accurve_prei1 <- specaccum(post_inoc_pre_stress_i1, method = "random", permutations = 400)
plot(accurve_prei1$sites, accurve_prei1$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc pre-stress ASVs - Inoculation 1")

accurve_prei2 <- specaccum(post_inoc_pre_stress_i2, method = "random", permutations = 400)
plot(accurve_prei2$sites, accurve_prei2$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc pre-stress ASVs - Inoculation 2")

accurve_prei3 <- specaccum(post_inoc_pre_stress_i3, method = "random", permutations = 400)
plot(accurve_prei3$sites, accurve_prei3$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc pre-stress ASVs - Inoculation 3")

accurve_prei4 <- specaccum(post_inoc_pre_stress_i4, method = "random", permutations = 400)
plot(accurve_prei4$sites, accurve_prei4$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc pre-stress ASVs - Inoculation 4")

accurve_prei5 <- specaccum(post_inoc_pre_stress_i5, method = "random", permutations = 400)
plot(accurve_prei5$sites, accurve_prei5$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc pre-stress ASVs - Inoculation 5")

accurve_prei6 <- specaccum(post_inoc_pre_stress_i6, method = "random", permutations = 400)
plot(accurve_prei6$sites, accurve_prei6$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc pre-stress ASVs - Inoculation 6")


##POST INOC POST STRESS #####################################################

accurve_posti1w1 <- specaccum(post_inoc_post_stress_i1_w1, method = "random", permutations = 400)
plot(accurve_posti1w1$sites, accurve_posti1w1$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i1 x w1")

accurve_posti1w2 <- specaccum(post_inoc_post_stress_i1_w2, method = "random", permutations = 400)
plot(accurve_posti1w2$sites, accurve_posti1w2$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i1 x w2")

accurve_posti1w3 <- specaccum(post_inoc_post_stress_i1_w3, method = "random", permutations = 400)
plot(accurve_posti1w3$sites, accurve_posti1w3$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i1 x w3")

accurve_posti2w1 <- specaccum(post_inoc_post_stress_i2_w1, method = "random", permutations = 400)
plot(accurve_posti2w1$sites, accurve_posti2w1$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i2 x w1")

accurve_posti2w2 <- specaccum(post_inoc_post_stress_i2_w2, method = "random", permutations = 400)
plot(accurve_posti2w2$sites, accurve_posti2w2$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i2 x w2")

accurve_posti2w3 <- specaccum(post_inoc_post_stress_i2_w3, method = "random", permutations = 400)
plot(accurve_posti2w3$sites, accurve_posti2w3$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i2 x w3")

accurve_posti3w1 <- specaccum(post_inoc_post_stress_i3_w1, method = "random", permutations = 400)
plot(accurve_posti3w1$sites, accurve_posti3w1$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i3 x w1")

accurve_posti3w2 <- specaccum(post_inoc_post_stress_i3_w2, method = "random", permutations = 400)
plot(accurve_posti3w2$sites, accurve_posti3w2$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i3 x w2")

accurve_posti3w3 <- specaccum(post_inoc_post_stress_i3_w3, method = "random", permutations = 400)
plot(accurve_posti3w3$sites, accurve_posti3w3$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i3 x w3")

accurve_posti4w1 <- specaccum(post_inoc_post_stress_i4_w1, method = "random", permutations = 400)
plot(accurve_posti4w1$sites, accurve_posti4w1$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i4 x w1")

accurve_posti4w2 <- specaccum(post_inoc_post_stress_i4_w2, method = "random", permutations = 400)
plot(accurve_posti4w2$sites, accurve_posti4w2$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress - i4 x w2")

accurve_posti4w3 <- specaccum(post_inoc_post_stress_i4_w3, method = "random", permutations = 400)
plot(accurve_posti4w3$sites, accurve_posti4w3$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress - i4 x w3")

accurve_posti5w1 <- specaccum(post_inoc_post_stress_i5_w1, method = "random", permutations = 400)
plot(accurve_posti5w1$sites, accurve_posti5w1$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i5 x w1")

accurve_posti5w2 <- specaccum(post_inoc_post_stress_i5_w2, method = "random", permutations = 400)
plot(accurve_posti5w2$sites, accurve_posti5w2$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i5 x w2")

accurve_posti5w3 <- specaccum(post_inoc_post_stress_i5_w3, method = "random", permutations = 400)
plot(accurve_posti5w3$sites, accurve_posti5w3$richness,
     xlab = "Number of sample",
     ylab = "Species richness",
     main = "Post-inoc post-stress ASVs - i5 x w3")

```
```{r, alpha diversity, eval=FALSE}

#############################################################################
#####  POST-DECONTAM: SEQ DEPTH CURVES (ALPHA DIVERSITY)  ###################
#############################################################################

par(mfrow=c(4,5), mar=c(1,1,1,1))

##POST INOC PRE STRESS#######################################################

#post_inoc_pre_stress_i1
rarecurve(post_inoc_pre_stress_i1,
          step = 1, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc pre-stress ASVs - Inoculation 1")

#post_inoc_pre_stress_i2
rarecurve(post_inoc_pre_stress_i2, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc pre-stress ASVs - Inoculation 2")

#post_inoc_pre_stress_i3
rarecurve(post_inoc_pre_stress_i3, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size",
          ylab = "ASV", 
          main = "Post-inoc pre-stress ASVs - Inoculation 3")

#post_inoc_pre_stress_i4
rarecurve(post_inoc_pre_stress_i4, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc pre-stress ASVs - Inoculation 4")

#post_inoc_pre_stress_i5
rarecurve(post_inoc_pre_stress_i5, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc pre-stress ASVs - Inoculation 5")

#post_inoc_pre_stress_i6
rarecurve(post_inoc_pre_stress_i6, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc pre-stress ASVs - Inoculation 6")

##POST INOC POST STRESS#######################################################

##INOCULATION 1
#post_inoc_post_stress_i1_w1
rarecurve(post_inoc_post_stress_i1_w1, 
          step = 20,
          label = FALSE,
          col = "blue", 
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i1 x w1")

#post_inoc_post_stress_i1_w2
rarecurve(post_inoc_post_stress_i1_w2, 
          step = 20, 
          col = "blue",
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i1 x w2")

#post_inoc_post_stress_i1_w3
rarecurve(post_inoc_post_stress_i1_w3, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i1 x w3")

##INOCULATION 2
#post_inoc_post_stress_i2_w1
rarecurve(post_inoc_post_stress_i2_w1, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i2 x w1")

#post_inoc_post_stress_i2_w2
rarecurve(post_inoc_post_stress_i2_w2, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i2 x w2")

#post_inoc_post_stress_i2_w3
rarecurve(post_inoc_post_stress_i2_w3, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i2 x w3")

##INOCULATION 3
#post_inoc_post_stress_i3_w1
rarecurve(post_inoc_post_stress_i3_w1, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i3 x w1")

#post_inoc_post_stress_i3_w2
rarecurve(post_inoc_post_stress_i3_w2, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i3 x w2")

#post_inoc_post_stress_i3_w3
rarecurve(post_inoc_post_stress_i3_w3, 
          step = 20, 
          col = "blue", 
          cex = 0.6, 
          label = FALSE,
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i3 x w3")

##INOCULATION 4
#post_inoc_post_stress_i4_w1
rarecurve(post_inoc_post_stress_i4_w1, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i4 x w1")

#post_inoc_post_stress_i4_w2
rarecurve(post_inoc_post_stress_i4_w2, 
          step = 20, 
          col = "blue", 
          cex = 0.6, 
          label = FALSE,
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i4 x w2")

#post_inoc_post_stress_i4_w3
rarecurve(post_inoc_post_stress_i4_w3, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i4 x w3")


##INOCULATION 5 
#post_inoc_post_stress_i5_w1
rarecurve(post_inoc_post_stress_i5_w1, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i5 x w1")

#post_inoc_post_stress_i5_w2
rarecurve(post_inoc_post_stress_i5_w2, 
          step = 20, 
          col = "blue", 
          cex = 0.6, 
          label = FALSE,
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i5 x w2")

#post_inoc_post_stress_i5_w3
rarecurve(post_inoc_post_stress_i5_w3, 
          step = 20, 
          col = "blue", 
          label = FALSE,
          cex = 0.6, 
          xlab = "Sample Size", 
          ylab = "ASV", 
          main = "Post-inoc post-stress ASVs - i5 x w3")

```

subset the decontam dataframe
```{r, subset phyloseq}
# subset phyloseq, now decontam'd

########## pre stress samples
post_inoc_pre_stress_ps.noncontam <- subset_samples(ps.noncontam,
                          sample_data(ps.noncontam)$SampleType=="post_inoc_pre_stress")

# remove ASVs not present in >1 sample
# 248 remain
post_inoc_pre_stress_ps.noncontam<- 
  prune_taxa(taxa_sums(post_inoc_pre_stress_ps.noncontam) > 1, post_inoc_pre_stress_ps.noncontam)

# estimate richness
rich_pre.stress<-as.data.frame(estimate_richness(post_inoc_pre_stress_ps.noncontam,
                                                 measure="Observed"))

#pull out tax table and metadata from post_inoc_PRE_stress to create new phyloseq object
pipre_metadata <- as(sample_data(post_inoc_pre_stress_ps.noncontam), 'matrix')
pipre_metadata <- as.data.frame(pipre_metadata)
pipre_taxtable <- as(tax_table(post_inoc_pre_stress_ps.noncontam), 'matrix')

# bind together
pipre_metadata <- dplyr::bind_cols(pipre_metadata, rich_pre.stress)

# plot richness
rich.pre.plot<- ggplot(pipre_metadata, 
              aes(x=Inoculation, y=Observed, fill=Water)) + ylab("Observed Richness") +
              geom_boxplot() + theme_light() + facet_wrap(~ Type)

rich.pre.plot
dev.copy(pdf, "figures/pre.stress.richness.pdf", height=4, width=6)
dev.off() 




########### ########### 
########### post stress samples
post_inoc_post_stress_ps.noncontam <- subset_samples(ps.noncontam,
                          sample_data(ps.noncontam)$SampleType=="post_inoc_post_stress")

# remove ASVs not present in >1 sample
# 1798 remain
post_inoc_post_stress_ps.noncontam<-
  prune_taxa(taxa_sums(post_inoc_post_stress_ps.noncontam) > 1, post_inoc_post_stress_ps.noncontam)

# estimate richness
rich_post.stress<-as.data.frame(estimate_richness(post_inoc_post_stress_ps.noncontam,
                                                  measure="Observed"))

#pull out tax table and metadata from post_inoc_PRE_stress to create new phyloseq object
pips_metadata <- as(sample_data(post_inoc_post_stress_ps.noncontam), 'matrix')
pips_metadata <- as.data.frame(pips_metadata)
pips_taxtable <- as(tax_table(post_inoc_post_stress_ps.noncontam), 'matrix')

# bind together
pips_metadata <- dplyr::bind_cols(pips_metadata, rich_post.stress)

# plot richness
rich.post.plot<- ggplot(pips_metadata, 
              aes(x=Inoculation, y=Observed, fill=Water)) + ylab("Observed Richness") +
              geom_boxplot() + theme_light() + facet_wrap(~ Type)


rich.post.plot
dev.copy(pdf, "figures/post.stress.richness.pdf", height=4, width=8)
dev.off() 


```

**NMDS and Permanova**
This data explores the impact of stress on the fungal communities.  
First, drop all the post-inoc/pre-stress samples from future analysis. Then Hellinger-transform the ASVs in an ASV contingency table. Using the new table, calculate the relative abundance for ASVs by summing the new totals for all ASVs (the transformed number of total reads) within a replicate and then dividing the transformed number of reads for each ASV by this total.

``` {r, NMDS permanova}
#############################################################################
################  POST-DECONTAM: NMDS  ######################################
#############################################################################

library(labdsv)
library(metagMisc)

#determine relative abundance + NMDS plot - drop all the post-inoc/pre-stress samples from future analysis

#post_inoc_post_stress table 
post_inoc_post_stress_asv <- as.data.frame(otu_table(post_inoc_post_stress_ps.noncontam))

#Hellinger transformation for all replicates in post-inoculation/post-stress ASV contingency table
transformed_pips_ASV <- hellinger(post_inoc_post_stress_asv)

#On this transformed data, calculate a “relative abundance” by summing the new totals for all ASVs (the transformed number of total reads) within a replicate and then dividing the transformed number of reads for each ASV by this total
rel_abund <- transformed_pips_ASV / rowSums(transformed_pips_ASV) 

#Create new phyloseq object with transformed table
otu <- otu_table(rel_abund,taxa_are_rows = FALSE)
tax <- tax_table(pips_taxtable)
met <- sample_data(pips_metadata)
row.names(met) <- pips_metadata$SampleID

ps.pips.hell <- phyloseq(otu,met,tax)
```

NMDS plot.

Using this new table of hellinger transformed ASVs, calculate Bray-Cutis community dissimilarities for all replicates and build an NMDS visualization where you code inoculum type by color and water treatment by shape.  

``` {r, NMDS plot}
GP.ord <- ordinate(ps.pips.hell, "NMDS", "bray")
NMDS.pips1 = plot_ordination(ps.pips.hell, GP.ord, type="samples", color="Inoculation", shape = "Water", title="Bray-Cutis community dissimilarities for all replicates")

print(NMDS.pips1)
dev.copy(pdf, "figures/post.stress.NMDS.pdf", height=5, width=6)
dev.off() 


```


``` {r, permanova}
## PERMANOVA #################################################################
library(vegan)

#bray curtis distance matrix
ps.pips.hell.bray <- phyloseq::distance(ps.pips.hell, method = "bray")

#make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps.pips.hell))

#---------
#run PERMANOVA for inoculation and water
# with interaction 
adonis_combo_int = adonis2(ps.pips.hell.bray ~ Inoculation * Water, permutations = 9999, 
                       data = sampledf, method = "bray")
print(adonis_combo_int)

# Homogeneity of dispersion test
beta.Inoc <- betadisper(ps.pips.hell.bray, sampledf$Inoculation)
permutest(beta.Inoc)

#This model shows all levels are significant, but inoculation is most important at R2 = 23%
```




#### start here : outlier removal makes no sense
```{r, anova}
#############################################################################
################  POST-DECONTAM: ANOVA  #####################################
#############################################################################

library(tidyverse)
library(car)
library(ggpubr)
library(effects)

dat = read.csv("data/pips_data/pips_complete.csv") %>%
  mutate(dat,
    Inoculation = as.character(Inoculation),
    Water = as.character(Water)
  )

ggplot(dat, aes(Inoculation, SpeciesRichness, fill = Water)) +
  facet_wrap(~ Type) +
  #scale_y_log10() +
  geom_boxplot()


## INITIAL EXPLORATORY ANALYSIS TO DETERMINE OUTLIERS - SPECIES RICHNESS: ROOTS AND SHOOTS ########################################
# interaction model
fit1 = lm(log(SpeciesRichness) ~ Type + Inoculation * Water, data = dat)

# additive model
fit2 = lm(log(SpeciesRichness) ~ Type + Inoculation + Water, data = dat)

anova(fit1, fit2)
AIC(fit1, fit2)
# fit 2 superior
Anova(fit2, type=2)

#plot to check residuals
plot(fit1)
plot(fit2)

#remove the outliers, re-run the analysis
filtered_dat = dat[-c(401, 78, 34, 81, 123), ]

#interaction model
filtered_fit1 = lm(log(SpeciesRichness) ~ Type + Inoculation * Water, data = filtered_dat)

#additive model 
filtered_fit2 = lm(log(SpeciesRichness) ~ Type + Inoculation + Water, data = filtered_dat)
plot(filtered_fit2)

anova(filtered_fit1, filtered_fit2)
AIC(filtered_fit1, filtered_fit2)

# models about the same, but interaction meaningless
Anova(filtered_fit2, type=2)

#check the outliers again
shapiro.test(filtered_fit1$residuals)
leveneTest(log(SpeciesRichness) ~ Water, data = filtered_dat)

filtered_dat <- filtered_dat %>%
  mutate(filtered_dat, res = resid(filtered_fit1))

#plot the filtered data to check the spread
ggplot(filtered_dat, aes(log(SpeciesRichness))) +
  facet_grid(Water ~ Inoculation) +
  # scale_x_log10() +
  geom_histogram()

#Re-plot grouped boxplot by inoculation and plant part without the outliers
ggplot(filtered_dat, aes(x=Inoculation, y=SpeciesRichness, fill=Water)) + 
    geom_boxplot() + theme_light() + facet_wrap(~ Type)
```


```{r, anova roots}
## SPECIES RICHNESS: ROOTS ###################################################

root_dat = subset(filtered_dat, Type == "Root")

#interaction model
root_fit1 = aov(log(SpeciesRichness) ~ Inoculation * Water, data = root_dat)
Anova(root_fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(root_fit1)
plot(allEffects(root_fit1))

#additive model 
root_fit2 = aov(log(SpeciesRichness) ~ Inoculation + Water, data = root_dat)
Anova(root_fit2, type = "2")
plot(root_fit2)
plot(allEffects(root_fit2))

## SPECIES RICHNESS: SHOOTS  #################################################

shoot_dat = subset(filtered_dat, Type == "Shoot")

#interaction model
shoot_fit1 = aov(log(SpeciesRichness) ~ Inoculation * Water, data = shoot_dat)
Anova(shoot_fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(shoot_fit1)
plot(allEffects(shoot_fit1))

#additive model 
shoot_fit2 = aov(log(SpeciesRichness) ~ Inoculation + Water, data = shoot_dat)
Anova(shoot_fit2, type = "2")
plot(shoot_fit2)
plot(allEffects(shoot_fit2))


## TOTAL BIOMASS: ROOTS  #####################################################

#interaction model
root_mass1= aov(WeightGrams ~ Inoculation * Water, data = root_dat)
Anova(root_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(root_mass1)
plot(allEffects(root_mass1))


#additive model
root_mass2 = aov(WeightGrams ~ Inoculation + Water, data = root_dat)
Anova(root_mass2, type = "2")
plot(root_mass2)
plot(allEffects(root_mass2))

## TOTAL BIOMASS: SHOOTS #####################################################
#interaction model
shoot_mass1 = aov(WeightGrams ~ Inoculation * Water, data = shoot_dat)
Anova(shoot_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(shoot_mass1)
plot(allEffects(shoot_mass1))

#additive model
shoot_mass2 = aov(WeightGrams ~ Inoculation + Water, data = shoot_dat)
Anova(shoot_mass2, type = "2")
plot(shoot_mass2)
plot(allEffects(shoot_mass2))


## ROOT:SHOOT RATIO ##########################################################

#load data with root:shoot ratio and total biomass (values calculated in excel)
rs_dat = read_csv("data/pips_data/pips_rootshoot_ratios.csv") |>
  mutate(rs_dat,
    Inoculation = as.character(Inoculation),
    Water = as.character(Water),
  )

ggplot(rs_dat, aes(Inoculation, RootShootRatio, fill = Water)) +
  scale_y_log10() +
  geom_boxplot()

#interaction model
rs_fit1 = aov(RootShootRatio ~ Inoculation * Water, data = rs_dat)
Anova(rs_fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(rs_fit1)

#additive model 
rs_fit2 = aov(RootShootRatio ~ Inoculation + Water, data = rs_dat)
Anova(rs_fit2, type = "2")
plot(rs_fit2)

# Studentized residuals
MASS::studres(rs_fit1) |>
  abs() |>
  sort()

#remove the outliers, re-run the analysis
filtered_rs = rs_dat[-c(49), ]

#plot the filtered data to check the spread
ggplot(filtered_rs, aes(RootShootRatio)) +
  facet_grid(Water ~ Inoculation) +
  # scale_x_log10() +
  geom_histogram()

#interaction model
rs_fit1 = aov(log(RootShootRatio) ~ Inoculation * Water, data = filtered_rs)
Anova(rs_fit1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(rs_fit1)

shapiro.test(rs_fit1$residuals)

#additive model 
rs_fit2 = aov(log(RootShootRatio) ~ Inoculation + Water, data = filtered_rs)
Anova(rs_fit2, type = "2")
plot(rs_fit2)

shapiro.test(rs_fit2$residuals)


rootmass = kruskal.test(WeightGrams ~ Inoculation, data = root_dat)
print(rootmass)

shootmass =  kruskal.test(WeightGrams ~ Inoculation + Water, data = shoot_dat)

```


```{r}
## TOTAL BIOMASS: ROOTS  #####################################################

#interaction model
root_mass1= aov(WeightGrams ~ Inoculation * Water, data = root_dat)
Anova(root_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(root_mass1)
plot(allEffects(root_mass1))

#additive model
root_mass2 = aov(WeightGrams ~ Inoculation + Water, data = root_dat)
Anova(root_mass2, type = "2")
plot(root_mass2)
plot(allEffects(root_mass2))

## TOTAL BIOMASS: SHOOTS #####################################################
#interaction model
shoot_mass1 = aov(WeightGrams ~ Inoculation * Water, data = shoot_dat)
Anova(shoot_mass1, type = "3") # if you have interactions, type-3 sums-of-squares is appropriate
plot(shoot_mass1)
plot(allEffects(shoot_mass1))

#additive model
shoot_mass2 = aov(WeightGrams ~ Inoculation + Water, data = shoot_dat)
Anova(shoot_mass2, type = "2")
plot(shoot_mass2)
plot(allEffects(shoot_mass2))

#Re-plot grouped boxplot by inoculation and plant part without the outliers
ggplot(filtered_dat, aes(x=Inoculation, y=WeightGrams, fill=Water)) + 
    geom_boxplot() + theme_light() + facet_wrap(~ Type)

```
